---
title: "Dyngo-4a Treatment Split Retina"
date: "2024-01-09"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In Melanoma-Associated Retinopathy (MAR), retinal rod bipolar cells internalize anti-TRPM1 autoantibodies, bringing them into their cytosol using an unknown mechanism. This experiment serves to elucidate the cellular mechanisms contributing to antibody uptake by pharmacologically inhibiting various endocytic pathways and observing the consequences on antibody uptake as seen in an *ex vivo* split mouse retina.

In this project, the effect of inhibiting dynamin-dependent endocytosis is observed by treating split retinas with Dyngo-4a. Briefly, the photoreceptors were removed from freshly isolated mouse retinas which were then incubated with an antibody for 1hr in the presence or absence of Dyngo 4a. The retinas were washed, then incubated with a viability dye to track cell death. Afterward, retinas were fixed, and the localization of the internalized antibodies was revealed by incubating the tissue with a labeled secondary antibody that targets the internalized antibody. 

Microscopy data were collected on a Leica SPX8 Confocal microscope. Images were processed in Arivis software using a custom pipeline that segments true uptake signal from internalized antibody based on a machine learning algorithm trained on these data. This segmentation is trained to ignore the super bright signal produced from dead cells and staining artifacts. ON-BCs are counted using the Cellpose algorithm, and average uptake is calculated by summing the uptake signal and dividing by the number of ON-BCs. 

Channel guide:

  - Channel 1 = Zombie dye
  - Channel 2 = Td tomato
  - Channel 3 = AF-488 (uptake)
  - Channel 4 = DIC

# Data Import and Tidying
```{r, include=FALSE}
library(tidyverse)
library(janitor)
library(readxl)
library(here)
library(kableExtra)
library(ggpubr)
library(rstatix)
source("uptake_functions.R")
```
**Pipeline Version**: This document has been updated to analyze data processed by Arivis pipeline:  "Ab-uptake_split-retina_ML_V1.pipeline"


## Read in the Data

For each individual experiment, there is a separate excel workbook of data generated for every image. These are workbooks are placed into subfolders of the *data* directory. All workbooks from an experiment are imported into a list, then merged into a single data frame with `bind_rows`

Variable names are then cleaned with `janitor` and extraneous columns are excluded (can be retrieved for different analysis later)

The `read_n_clean` function also implements the `condition_select` method, which allows the user to assign treatment conditions (i.e DMSO vs Dyngo-4a) to samples by the name of the sample (i.e "s1.2)
```{r}
#e1-449
DMSO_imgs <- c("s2.1","s2.3") #removed s1.2
dyngo_imgs <- c("s3.1", "s3.2", "s3.3", "s4.1", "s4.3")
e449 <- read_n_clean(folder_path = "data/e1-449", cond1_vec = DMSO_imgs, cond2_vec = dyngo_imgs, cond1_title = "DMSO", cond2_title = "Dyngo-4a")

#e1-455
DMSO_imgs <- c("s1.1", "s1.3", "s2.1")
dyngo_imgs <- c("s3.2", "s4.1")
e455 <- read_n_clean(folder_path = "data/e1-455", cond1_vec = DMSO_imgs, cond2_vec = dyngo_imgs, cond1_title = "DMSO", cond2_title = "Dyngo-4a")

#e1-456
DMSO_imgs <- c("s2.1", "s2.2", "s2.3") #potentially remove s2.1
dyngo_imgs <- c("s3.3")
e456 <- read_n_clean(folder_path = "data/e1-456", cond1_vec = DMSO_imgs, cond2_vec = dyngo_imgs, cond1_title = "DMSO", cond2_title = "Dyngo-4a")

#e1-458
DMSO_imgs <- c("none") #only dyngo images for this data set
dyngo_imgs <- c("s3.2")
e458 <- read_n_clean(folder_path = "data/e1-458", cond1_vec = DMSO_imgs, cond2_vec = dyngo_imgs, cond1_title = "DMSO", cond2_title = "Dyngo-4a")

#place all experiment dataframes into a list for batch processing
all_dyngo_exps <- list("e449" = e449, "e455" = e455, "e456" = e456, "e458" = e458)
```


# Data Analysis

The current analysis pipeline produces two object types: "Segment ON-BCs" and "Segment Uptake." The number of Segment ON-BC objects equals the number of ON-BCs counted in the image. The number of Segment Uptake objects is arbitrary, but represent the segmentation objects identified by the ML model. To get the total uptake signal, the "sum_intensities_3" value is summed for every one of the Segment Uptake objects. The value "avergae uptake per cell" is calculated by dividing this sum by the average number of bipolar cells counted across all images.
<p style="text-align: center;">s1.2<span style="color:blue">**-1**</span></p>

Begin by identifying all images that come from the same piece of tissue, using a Regex to identify this based on the sample name
```{r}
#regex string that will extract everything between the 's' and the '-' in the sample name
sample_string <- "s.*?.?(?=-)"

#add new variable "tissue" that uses the regex to match all images from the same piece of tissue
all_dyngo_exps <- lapply(all_dyngo_exps, function(x){mutate(x, tissue = str_extract(x$image_set, pattern = sample_string),.after = 1)})
all_dyngo_exps[1]
```
## Calculate the average uptake per ON-BC (`avg_upc`)

`sum_intensities_5` represents the total amount of 488-uptake signal in an image (after filtering). This will be divided by the total number of ON-BCs to produce `avg_upc`

- images with low Td-tomato signal intensity are not being counted properly. To balance inacurate counts, the average number of ON-BCs across the entire data set will be calculated after merging all data frames into one

- total number of ON-BCs can be found by counting rows with ON-BCs in the name
- as of pipeline V5.4, Channel 5 channel in the image is a new channel that stores the result of applying an intensity filter threshold to the 488 channel (Channel 3).

```{r}
all_dyngo_exps <- lapply(all_dyngo_exps, avg_uptake)
```

Merge all dataframes in list into one dataframe
```{r}
all_dyngo_exps <- bind_rows(all_dyngo_exps, .id = "exp") %>% 
  mutate(exp = as.factor(exp))

#averaging the number of ON-BCs across all images in data set to account for dim Td tom signal that caused miscount
all_dyngo_exps["avg_on_bc_count"] <- mean(all_dyngo_exps$on_bc_count) 

#average uptake per cell, sum intensity for a single image over the avg ON-BC count for all experiments
all_dyngo_exps["avg_upc"] <- all_dyngo_exps$sum_intensities/all_dyngo_exps$avg_on_bc_count 
```


## Averaging across tissue
Next, group by `exp` and  `tissue`, then average `avg_upc`
```{r}
all_dyngo_exps <- all_dyngo_exps %>% 
    group_by(exp, tissue) %>% #first compute avg grouping by tissue (contains multiple images)
    mutate(tissue_avg = mean(avg_upc)) %>% 
    ungroup()
```

Checking to see that each image has a distinct average and each tissue has a distinct average. Looks good
```{r}
all_dyngo_exps %>%
  group_by(image_set, exp, tissue, condition, sum_intensities, avg_on_bc_count, avg_upc, tissue_avg) %>% 
  summarise() %>% 
  arrange(exp) %>% 
  kbl() %>% 
  kable_paper()
```


When plotting, each piece of tissue (tissue average) will be a point. The bar will be the *condition average* and will be made after merging all experiments into a single dataframe. 
```{r}
all_dyngo_exps <- all_dyngo_exps %>% 
    group_by(condition) %>% 
    mutate(condition_avg = mean(unique(tissue_avg)),
           condition_sd = sd(tissue_avg)) %>% 
    ungroup()
```

## Validation and testing

Verifying avg_upc calculation
```{r}
all_dyngo_exps %>% 
  group_by(image_set, tissue,exp, condition,sum_intensities, on_bc_count, avg_upc, tissue_avg) %>% 
  summarise() %>% 
  filter(exp == "e449") %>% 
  ggplot()+
  geom_col(aes(x = image_set, y = avg_upc))+
  geom_point(aes(x = image_set, y = tissue_avg), color = "orange")+
  #geom_col(aes(x = image_set, y = on_bc_count/on_bc_count[1]), fill = "darkgreen")
  theme(axis.text.x = element_text(angle = 90))
```
Making sure that image intensities (as seen by eye) actually match plotted values
```{r}
all_dyngo_exps %>% 
  group_by(image_set, tissue,exp, condition,sum_intensities, on_bc_count, avg_upc, tissue_avg) %>% 
  summarise() %>% 
  filter((exp == "e449" & tissue == "s2.1") | (exp == "e456" & tissue == "s2.3")) %>% #for selecting specific tissue within specific exp
  #filter(exp == "e456") %>% 
  ggplot()+
  #geom_col(aes(x = image_set, y = tissue_avg), fill = "orange")+
  geom_col(aes(x = image_set, y = sum_intensities, fill = condition))+
  #geom_col(aes(x = image_set, y = on_bc_count/on_bc_count[1]), fill = "darkgreen")
  theme(axis.text.x = element_text(angle = 90))

#ggsave("/Users/ryanhecht/Desktop/e1-449_e1-456_dmso.pdf")
```



This visual verifies that the tissue_avg is being calculated appropriately
# Statistics

Data includes one continuous outcome variable (uptake signal intensity), one categorical predictor variable (treatment condition). If the data are normally distributed, then significant difference between groups can be assessed using an `independent t test (unpaired)`

## Shapiro-Wilk Normality test

- performing on just what goes into the plot: `tissue_avg`

Histogram on points to be plotted
```{r}
all_dyngo_exps %>% 
  group_by(tissue, condition, condition_avg, tissue_avg, condition_sd, exp) %>%
  summarise() %>% 
  pull(tissue_avg) %>%
  hist()
```
Shapiro-Wilk on points to be plotted
```{r}
shapiro_result <- all_dyngo_exps %>% 
  group_by(tissue, condition, condition_avg, tissue_avg, condition_sd, exp) %>%
  summarise() %>% 
  pull(tissue_avg) %>% 
  shapiro.test()

if(shapiro_result$p.value > 0.05){
  print(c("The p-value of the Shapiro-Wilks test = ", as.character(shapiro_result$p.value), ". For alpha = 0.05, the data are normally distributed"))
}
if(shapiro_result$p.value < 0.05){
  print(c("The p-value of the Shapiro-Wilks test = ", as.character(shapiro_result$p.value), ". For alpha = 0.05, the data are NOT normally distributed"))
}

```
Visually, the data appears very close to normally distributed.


## Un-paired T-test

Performing a Welch's t-test since this does not assume the variance in the two groups is the same.
```{r}
dyngo_t_test <- all_dyngo_exps %>% 
  group_by(tissue, condition, condition_avg, tissue_avg, condition_sd, exp) %>%
  summarise() %>% #the group by and summarize lines are to filter to only the data that will be plotted and was tested in the normality test
  ungroup() %>% 
  t_test(tissue_avg ~ condition) %>% 
  add_significance() %>% 
  add_xy_position(x = "condition")

dyngo_t_test
```

# Plotting

```{r, fig.height=7, fig.width=6}
dyngo_plot <- uptake_plot(all_dyngo_exps, output_name = "all_dyngo_exps_V5.4", condition_label = "Condition", add_stats = TRUE, dyngo_t_test)
dyngo_plot
```
Table of the plotted values
```{r}
all_dyngo_exps %>% 
  group_by(tissue, condition, condition_avg, tissue_avg, condition_sd, exp) %>%
  summarise() %>% 
  arrange(condition, tissue_avg) %>% 
  kbl() %>% 
  kable_paper()
```


- **NOTE** the pipeline now corrects for coverslip reflections, as any images with reflections have been cropped out. So this should be considered the true uptake signal

Plotting tissue average by condition and experiment number
```{r}
all_dyngo_exps %>% 
  ggplot()+
  geom_point(aes(x = exp, y = tissue_avg, color = condition), alpha = 0.7)+
  theme_classic()
```



