---
title: "Imipramine Treatment Split Retina"
date: "2024-03-29"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In Melanoma-Associated Retinopathy (MAR), retinal rod bipolar cells internalize anti-TRPM1 autoantibodies, bringing them into their cytosol using an unknown mechanism. This experiment serves to elucidate the cellular mechanisms contributing to antibody uptake by pharmacologically inhibiting various endocytic pathways and observing the consequences on antibody uptake as seen in an *ex vivo* split mouse retina.

In this project, the effect of inhibiting dynamin-dependent endocytosis is observed by treating split retinas with imipramine-4a. Briefly, the photoreceptors were removed from freshly isolated mouse retinas which were then incubated with an antibody for 1hr in the presence or absence of imipramine 4a. The retinas were washed, then incubated with a viability dye to track cell death. Afterward, retinas were fixed, and the localization of the internalized antibodies was revealed by incubating the tissue with a labeled secondary antibody that targets the internalized antibody. 

Microscopy data were collected on a Leica SPX8 Confocal microscope. Images were processed in Arivis software using a custom pipeline that quantifies internalized antibody only within rod bipolar cells (segmented based on transgenic Td tomato expression), excluding any dead cells or bright staining artefacts. 

Channel guide:

  - Channel 1 = Zombie dye
  - Channel 2 = Td tomato
  - Channel 3 = AF-488 (uptake)
  - Channel 4 = DIC

# Data Import and Tidying
```{r, include=FALSE}
library(tidyverse)
library(janitor)
library(readxl)
library(here)
library(kableExtra)
library(ggpubr)
library(rstatix)
source("uptake_functions.R")
```
**Pipeline Version**: this notebook is updated to process output from Arivis pipeline V4.2


## Read in the Data

For each individual experiment, there is a separate excel workbook of data generated for every image. These are workbooks are placed into subfolders of the *data* directory. All workbooks from an experiment are imported into a list, then merged into a single data frame with `bind_rows`

Variable names are then cleaned with `janitor` and extraneous columns are excluded (can be retrieved for different analysis later)

The `read_n_clean` function also implements the `condition_select` method, which allows the user to assign treatment conditions (i.e DMSO vs imipramine) to samples by the name of the sample (i.e "s1.2")
```{r}
#e1-459
# DMSO_imgs <- c("s1.1", "s1.2")
# imipramine_imgs <- c("s3.1", "s3.2", "s3.3")
# e459 <- read_n_clean(folder_path = "data/e1-459", cond1_vec = DMSO_imgs, cond2_vec = imipramine_imgs, cond1_title = "DMSO", cond2_title = "Imipramine")

#e1-460
DMSO_imgs <- c("s2.1", "s2.2", "s2.3")
imipramine_imgs <- c("s3.1", "s3.3", "s4.3")
e460 <- read_n_clean(folder_path = "data/e1-460", cond1_vec = DMSO_imgs, cond2_vec = imipramine_imgs, cond1_title = "DMSO", cond2_title = "Imipramine")

#e1-461
DMSO_imgs <- c("s2.2", "s2.3")
imipramine_imgs <- c("s4.3")
e461 <- read_n_clean(folder_path = "data/e1-461", cond1_vec = DMSO_imgs, cond2_vec = imipramine_imgs, cond1_title = "DMSO", cond2_title = "Imipramine")

#place all experiment dataframes into a list for batch processing
all_imipramine_exps <- list("e460" = e460, "e461" = e461)
```


# Data Analysis

These data are processed by a different Arivis pipeline than the Dyngo-4a and 4 ºC experiments owing to the fact that the repair of the Leica WLL increased the intensity of the pixel values while using the same acquisition settings. Since the enhanced brightness helped reveal more cell body uptake, I decided to leave the settings rather than adjust them to produce the same pixel values as the earlier experiments. The consequence is that there is cell body labeling associated with dead cells that falls below the intensity filter threshold which excluded "false uptake" or uptake from dead cells. This filter is still in place, and effectively removes signal caused by uptake in the *dendrites* of dead cells, but to ignore uptake from dead cell bodies, I moved back to an earlier version of the pipeline (V4.2) which only counts uptake that occurs within ON-BC cell bodies. I previously demonstrated that this strategy produces extremely similar results to the V5 pipeline.

The "sum_intensities_5" in the dataframe is the result of creating a new channel from channel 3 with the intensity threshold filter applied to remove fake uptake. This is the sum of pixel values that fall within any given ON-BC. There may be multiple (non-overlapping) images from a given piece of tissue, denoted by the number following the hyphen in the sample name:

<p style="text-align: center;">s1.2<span style="color:blue">**-1**</span></p>

Begin by identifying all images that come from the same piece of tissue, using a Regex to identify this based on the sample name
```{r}
#regex string that will extract everything between the 's' and the '-' in the sample name
sample_string <- "s.*?.?(?=-)"

#add new variable "tissue" that uses the regex to match all images from the same piece of tissue
all_imipramine_exps <- lapply(all_imipramine_exps, function(x){mutate(x, tissue = str_extract(x$image_set, pattern = sample_string),.after = 1)})
all_imipramine_exps[1]
```
## Calculate the average uptake per ON-BC (`avg_upc`)

Take `sum_intensities_5`, which represents the total amount of 488-uptake signal in a given "cell" (after filtering) and sum this value across all "cells" in an image. Then divide this by the total volume of "cells" in that image

- calculating by volume because the number of cells will not be accurate based on the way the blob finder algorithm is drawing cell boundaries
- Channel 5 channel in the image is a new channel that stores the result of applying an intensity filter threshold to the 488 channel (Channel 3).

```{r}
all_imipramine_exps <- lapply(all_imipramine_exps, avg_uptake_vol)
```


## Averaging across tissue
Next, group by `tissue` and average `avg_upc`
```{r}
all_imipramine_exps <- lapply(all_imipramine_exps, function(x){
  x %>%   
    group_by(tissue) %>% #first compute avg grouping by tissue (contains multiple images)
    mutate(tissue_avg = mean(unique(avg_upc))) %>% 
    ungroup()
  })

all_imipramine_exps[1]
```

Checking to see that each image has a distinct average and each tissue has a distinct average. Looks good
```{r}
all_imipramine_exps[[1]] %>%
    group_by(image_set, tissue, condition, avg_upc, tissue_avg) %>% 
    summarise() %>% 
  kbl() %>% 
  kable_paper()
```
Merge all dataframes in list into one dataframe
```{r}
all_imipramine_exps <- bind_rows(all_imipramine_exps, .id = "exp") %>% 
  mutate(exp = as.factor(exp))
```


When plotting, each piece of tissue (tissue average) will be a point. The bar will be the *condition average* and will be made after merging all experiments into a single dataframe. 
```{r}
all_imipramine_exps <- all_imipramine_exps %>% 
    group_by(condition) %>% 
    mutate(condition_avg = mean(unique(tissue_avg)),
           condition_sd = sd(tissue_avg)) %>% 
    ungroup()

#all_imipramine_exps
```

# Statistics

Data includes one continuous outcome variable (uptake signal intensity), one categorical predictor variable (treatment condition). If the data are normally distributed, then significant difference between groups can be assessed using an `independent t test (unpaired)`

## Shapiro-Wilk Normality test

- performing on just what goes into the plot: `tissue_avg`

Histogram on points to be plotted
```{r}
all_imipramine_exps %>% 
  group_by(tissue, condition, condition_avg, tissue_avg, condition_sd, exp) %>%
  summarise() %>% 
  pull(tissue_avg) %>%
  hist()
```
Shapiro-Wilk on points to be plotted
```{r}
shapiro_result <- all_imipramine_exps %>% 
  group_by(tissue, condition, condition_avg, tissue_avg, condition_sd, exp) %>%
  summarise() %>% 
  pull(tissue_avg) %>% 
  shapiro.test()

if(shapiro_result$p.value > 0.05){
  print(c("The p-value of the Shapiro-Wilks test = ", as.character(shapiro_result$p.value), ". For alpha = 0.05, the data are normally distributed"))
}
if(shapiro_result$p.value < 0.05){
  print(c("The p-value of the Shapiro-Wilks test = ", as.character(shapiro_result$p.value), ". For alpha = 0.05, the data are NOT normally distributed"))
}

```



## Un-paired T-test

Performing a Welch's t-test since this does not assume the variance in the two groups is the same.
```{r}
imipramine_t_test <- all_imipramine_exps %>% 
  group_by(tissue, condition, condition_avg, tissue_avg, condition_sd, exp) %>%
  summarise() %>% #the group by and summarize lines are to filter to only the data that will be plotted and was tested in the normality test
  ungroup() %>% 
  t_test(tissue_avg ~ condition) %>% 
  add_significance() %>% 
  add_xy_position(x = "condition")

imipramine_t_test
```

# Plotting

```{r, fig.height=7, fig.width=6}
imipramine_plot <- uptake_plot(all_imipramine_exps, output_name = "all_imipramine_exps_V5.4", condition_label = "Condition", add_stats = TRUE, imipramine_t_test)
imipramine_plot
```
Table of the plotted values
```{r}
all_imipramine_exps %>% 
  group_by(tissue, condition, condition_avg, tissue_avg, condition_sd, exp) %>%
  summarise() %>% 
  kbl() %>% 
  kable_paper()
```



Plotting tissue average by condition and experiment number
```{r}
all_imipramine_exps %>% 
  ggplot()+
  geom_point(aes(x = exp, y = tissue_avg, color = condition), alpha = 0.7)+
  theme_classic()
```


