---
title: "pitstop2 2"
output: html_document
date: "2024-07-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In Melanoma-Associated Retinopathy (MAR), retinal rod bipolar cells internalize anti-TRPM1 autoantibodies, bringing them into their cytosol using an unknown mechanism. This experiment serves to elucidate the cellular mechanisms contributing to antibody uptake by pharmacologically inhibiting various endocytic pathways and observing the consequences on antibody uptake as seen in an *ex vivo* split mouse retina.

In this project, the effect of inhibiting dynamin-dependent endocytosis is observed by treating split retinas with pitstop2-4a. Briefly, the photoreceptors were removed from freshly isolated mouse retinas which were then incubated with an antibody for 1hr in the presence or absence of pitstop2 4a. The retinas were washed, then incubated with a viability dye to track cell death. Afterward, retinas were fixed, and the localization of the internalized antibodies was revealed by incubating the tissue with a labeled secondary antibody that targets the internalized antibody. 

Microscopy data were collected on a Leica SPX8 Confocal microscope. Images were processed in Arivis software using a custom pipeline that segments true uptake signal from internalized antibody based on a machine learning algorithm trained on these data. This segmentation is trained to ignore the super bright signal produced from dead cells and staining artifacts. ON-BCs are counted using the Cellpose algorithm, and average uptake is calculated by summing the uptake signal and dividing by the number of ON-BCs. 

Channel guide:

  - Channel 1 = Zombie dye
  - Channel 2 = Td tomato
  - Channel 3 = AF-488 (uptake)
  - Channel 4 = DIC

# Data Import and Tidying
```{r, include=FALSE}
library(tidyverse)
library(janitor)
library(readxl)
library(here)
library(kableExtra)
library(ggpubr)
library(rstatix)
source("uptake_functions.R")
```
**Pipeline Version**: This document has been updated to analyze data processed by Arivis pipeline:  "Ab-uptake_split-retina_ML_V2.pipeline"
**Data processing date**: 24-07-15


## Read in the Data

For each individual experiment, there is a separate excel workbook of data generated for every image. These are workbooks are placed into subfolders of the *data* directory. All workbooks from an experiment are imported into a list, then merged into a single data frame with `bind_rows`

Variable names are then cleaned with `janitor` and extraneous columns are excluded (can be retrieved for different analysis later)

The `read_n_clean` function also implements the `condition_select` method, which allows the user to assign treatment conditions (i.e DMSO vs pitstop2-4a) to samples by the name of the sample (i.e "s1.2)
```{r}
#e1-449
DMSO_imgs <- c("s2.1","s2.3") #removed s1.2
pitstop2_imgs <- c("s3.1", "s3.2", "s3.3", "s4.1", "s4.3")
e449 <- read_n_clean(folder_path = "data/e1-449", cond1_vec = DMSO_imgs, cond2_vec = pitstop2_imgs, cond1_title = "DMSO", cond2_title = "pitstop2")

#e1-455
DMSO_imgs <- c("s1.1", "s1.3", "s2.1")
pitstop2_imgs <- c("s3.2", "s4.1")
e455 <- read_n_clean(folder_path = "data/e1-455", cond1_vec = DMSO_imgs, cond2_vec = pitstop2_imgs, cond1_title = "DMSO", cond2_title = "pitstop2")

#e1-456
DMSO_imgs <- c("s2.1", "s2.2", "s2.3") #potentially remove s2.1
pitstop2_imgs <- c("s3.3")
e456 <- read_n_clean(folder_path = "data/e1-456", cond1_vec = DMSO_imgs, cond2_vec = pitstop2_imgs, cond1_title = "DMSO", cond2_title = "pitstop2")

#e1-458
DMSO_imgs <- c("none") #only pitstop2 images for this data set
pitstop2_imgs <- c("s3.2")
e458 <- read_n_clean(folder_path = "data/e1-458", cond1_vec = DMSO_imgs, cond2_vec = pitstop2_imgs, cond1_title = "DMSO", cond2_title = "pitstop2")

#place all experiment dataframes into a list for batch processing
all_pitstop2_exps <- list("e449" = e449, "e455" = e455, "e456" = e456, "e458" = e458)
```
